Staged Abstract Interpreters

### 2019-09-11 paper/10_intro.tex
typo
misc

Roughly at the same time abstract interpretation was proposed,
\citet{futamura1971partial} observed a close connection between interpreters and
compilers through a hierarchy of specializations, known since then as the
Futamura Projections. The first Futamura projection states that specializing an
interpreter to an input program removes interpretative overhead and yields an
equivalent executable that runs faster. In recent years, the idea of Futamura
projections has attracted more attention, and has been successfully applied to
many real-world problems, including query compilers
\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc. The challenge here: can we
apply Futamura projections to a general abstract interpreter to make it run
faster? The solution is not obvious, as we need to specialize a nonstandard
semantics that extensively uses nondeterminism, operates on abstract domains,
and computes fixed-points iteratively over caches.

@@

Program specialization and abstract interpretation are two orthogonal
concepts.  To implement the confluence of them, we first construct a
generic interpreter that is agnostic to both binding times and value
domains used in the semantics.  Later, the generic interpreter can be
instantiated from these two dimensions (Figure~\ref{confluence}):
\begin{itemize}
\item With a flat binding-time and concrete domains, it is an ordinary
  definitional interpreter based on big-step operational semantics;
\item With two-level binding-times and concrete domains, it is a
  compiler that translates a program into another language;
\item With a flat binding-time and abstract domains, it is a
  definitional abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}
  that statically computes runtime properties;
\item With two-level binding-times and abstract domains, it is an optimizing
  program analyzer, but works in the fashion of compilation.
\end{itemize}

>>>>

Roughly at the same time abstract interpretation was proposed,
\citet{futamura1971partial} observed a close connection between interpreters and
compilers through a hierarchy of specializations, known since then as the
Futamura projections. The first Futamura projection states that specializing an
interpreter to an input program removes interpretative overhead and yields an
equivalent executable that runs faster. In recent years, the idea of Futamura
projections has attracted more attention, and has been successfully applied to
many real-world problems, including query compilers
\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc. The challenge here: can we
apply Futamura projections to a general abstract interpreter to make it run
faster? The solution is not obvious, as we need to specialize a nonstandard
semantics that extensively uses nondeterminism, operates on abstract domains,
and computes fixed-points iteratively over caches.

@@

Program specialization and abstract interpretation are two orthogonal
concepts.  To implement the confluence of them, we first construct a
generic interpreter that is agnostic to both binding times and value
domains used in the semantics.  Later, the generic interpreter can be
instantiated from these two dimensions (Figure~\ref{confluence}):
\begin{itemize}
\item With a flat binding-time and concrete domains, it is an ordinary
  definitional interpreter based on big-step operational semantics;
\item With two-level binding-times and concrete domains, it is a
  compiler that translates a program into another language;
\item With a flat binding-time and abstract domains, it is a
  definitional abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}
  that statically computes runtime properties;
\item With two-level binding-times and abstract domains, it is an optimized
  program analyzer, but works in the fashion of compilation.
\end{itemize}
